### 2020-10-15 conclusion.tex
done with related work

\begin{document}
\section{Related Work and Conclusion}
\textit{Simon: I think you really need some mention of MPSTs here, since a key motivation for MPSTs is ensuring deadlock-freedom between participants in a single MPST session. You can then talk about the global progress work by Coppo~\etal, too, which incorporates an 'interaction typing system' (that few people understand). How does that relate?}

\subsubsection*{Deadlock freedom}

Deadlock freedom and progress are well-studied in the $\pi$-calculus.
For the standard typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock-freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set. These abstract tags were later simplified to natural numbers. Pairs of obligations and capabilities were used in the type system for lock freedom~\cite{kobayashi02,kobayashi06}, which allowed more $\pi$-calculus processes to be typed. Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later simplified them to a single priority for linear $\pi$-calculus~\cite{padovani14}. Using the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16}, the priority-based technique for deadlock freedom can be transferred onto the $\pi$-calculus with session types.

For the session-typed $\pi$-calculus, the foundational work by Dezani~\etal~\cite{dezani-ciancaglinimostrous06} guarantees progress by allowing only one active session at a time. Dezani later~\cite{dezani-ciancagliniliguoro09progress} introduces a partial order on channels in line with Kobayashi's work~\cite{kobayashi98}. Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} used to provide a missing counterpart to a process, thus guaranteeing deadlock freedom.
Carbone~\etal~\cite{carbonedardha14} studied the use of catalysers and showed that progress is a compositional form of lock-freedom. Following Padovani~\cite{padovani14}, Dardha~\etal~\cite{dardhagiachino12} show that by using the encoding of session types and Kobayashi's obligations/capabilities, we can obtain progress for session types.\wen{I'm not sure what this means.} Vieira and Vasconcelos~\cite{vieiravasconcelos13} used single priorities and an abstract partial order to guarantee deadlock freedom in a session-typed $\pi$-calculus.

@@

\subsubsection*{Ties with logic}
The correspondence between logic and types lays the foundation for functional programming~\cite{wadler15}. Since its inception by Girard~\cite{girard87}, linear logic has been a candidate for a foundational correspondence for concurrent programs. A~correspondence with linear $\pi$-calculus was established early on by Abramsky~\cite{abramsky94,bellinscott94}. A~correspondence between session-typed $\pi$-calculus and dual intuitionistic linear logic was developed by Caires and Pfenning~\cite[$\pi\text{DILL}$]{cairespfenning10}, and with classical linear logic by Wadler~\cite[CP]{wadler15}, both guaranteeing deadlock freedom as a result of their connection to logic. Dardha and Gay~\cite[PCP]{dardhagay18} integrate Padovani's work on priorities~\cite{padovani14} with CP, creating the first calculus which combines priorities and strong ties with logic. Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and show that CLL corresponds to Kobayashi's system with the restriction that only single cuts are allowed.\wen{What do you mean by ``single cuts''?} Balzer~\etal~\cite[$\text{SILL}_S$]{balzerpfenning17} introduce shared state, which breaks deadlock freedom. They later restore deadlock freedom using priorities~\cite[$\text{SILL}_{S+}$]{balzertoninho19}.

>>>>

\begin{document}
\section{Related Work and Conclusion}
\subsubsection*{Deadlock freedom and progress}
Deadlock freedom and progress are well studied properties in the $\pi$-calculus.
For the standard typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set. These abstract tags were later simplified to natural numbers. Pairs of obligations and capabilities were used in the type system for deadlock freedom~\cite{kobayashi06}, which allowed more $\pi$-calculus processes to be typed. Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later simplified them to a single priority for linear $\pi$-calculus~\cite{padovani14}.  As discussed by the author, by combining the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16} and the priority-based linear $\pi$-calculus, we can obtain deadlock freedom for the $\pi$-calculus with session types.

For the session-typed $\pi$-calculus, the foundational work by Dezani~\etal~\cite{dezani-ciancaglinimostrous06} guarantees progress by allowing only one active session at a time. Dezani later~\cite{dezani-ciancagliniliguoro09progress} introduces a partial order on channels in line with Kobayashi's work~\cite{kobayashi98}. Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} used to provide a missing counterpart to a process, thus guaranteeing deadlock freedom.
Carbone~\etal~\cite{carbonedardha14} studied the use of catalysers and showed that progress is a compositional form of lock-freedom. In the same lines of Padovani~\cite{padovani14}, Carbone~\etal~\cite{carbonedardha14} show that progress can be lifted to session types via the encoding of session types to linear types and Kobayashi's obligations/capabilities. Vieira and Vasconcelos~\cite{vieiravasconcelos13} use single priorities and an abstract partial order to guarantee deadlock freedom in a session-typed $\pi$-calculus.

Multiparty Session Types (MPST) \cite{hondayoshida08} on the other hand guarantee deadlock freedom within a single session, but more advanced techniques are required for session interleaving. Bettini~\etal~\cite{bettinicoppo08} follow a similar technique to Kobayashi for multiparty session types. The main
difference with our work is that we associate priorities to communication actions, and not to channels. Carbone and Montesi~\cite{carbonemontesi13} explore and combine MPST and choreographies to obtain a formalism that satisfies deadlock freedom by design.

@@

\subsubsection*{Ties with logic}
The correspondence between logic and types lays the foundation for functional programming~\cite{wadler15}. Since its inception by Girard~\cite{girard87}, linear logic has been a candidate for a foundational correspondence for concurrent programs. A~correspondence with linear $\pi$-calculus was established early on by Abramsky~\cite{abramsky94,bellinscott94}. A~correspondence between session-typed $\pi$-calculus and dual intuitionistic linear logic was developed by Caires and Pfenning~\cite[$\pi\text{DILL}$]{cairespfenning10}, and with classical linear logic by Wadler~\cite[CP]{wadler15}, both guaranteeing deadlock freedom as a result of their connection to logic. Dardha and Gay~\cite[PCP]{dardhagay18} integrate Kobayashi's and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, creating the first calculus which combines priorities and strong ties with logic. Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and show that CLL corresponds to a subsystem of Kobayashi's where restriction is applied once. Balzer~\etal~\cite[$\text{SILL}_S$]{balzerpfenning17} introduce shared state, which breaks deadlock freedom. They later restore deadlock freedom using priorities~\cite[$\text{SILL}_{S+}$]{balzertoninho19}.
Carbone~\etal~\cite{carbonelindley16} give a logical view of MPST with a generalised duality, called coherence.
